FlowFPX: Nimble Tools for Debugging Floating-Point Exceptions

### 2023-06-16 paper.tex
Short draft of fp primer

\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\FlowFPX}{FlowFPX}
\newcommand{\GPUFPX}{GPU-FPX}
\newcommand{\FloatTracker}{FloatTracker}
\newcommand{\FT}{\FloatTracker}
\newcommand{\fp}{floating-point} % hyphen or no?
\newcommand{\CSTG}{stack graph}
\newcommand{\CPP}{\code{C++}}
\newcommand{\Dendro}{\textsc{Dendro}}
\newcommand{\urlaccess}[2]{\url{#1}}
\newcommand{\Nan}{\code{NaN}}
\newcommand{\NaN}{\Nan}
\newcommand{\Inf}{\code{Inf}}

@@

>>>>

\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\FlowFPX}{FlowFPX}
\newcommand{\GPUFPX}{GPU-FPX}
\newcommand{\FloatTracker}{FloatTracker}
\newcommand{\FT}{\FloatTracker}
\newcommand{\Fp}{Floating-point} % hyphen or no?
\newcommand{\fp}{floating-point} % hyphen or no?
\newcommand{\CSTG}{stack graph}
\newcommand{\CPP}{\code{C++}}
\newcommand{\Dendro}{\textsc{Dendro}}
\newcommand{\urlaccess}[2]{\url{#1}}
\newcommand{\Nan}{\code{NaN}}
\newcommand{\NaN}{\Nan}
\newcommand{\Inf}{\code{Inf}}

@@

\Fp{} numbers are a way of representing real numbers in a computer.
However, since we limit our precision to a fixed, finite set of bits, \fp{} is \emph{inherently approximate} and necessarily incurs some error during computation.
This is not bad; it is a trade-off: we get fast and compact representations that are good enough for most use cases.
However, the downside is that \fp{} numbers \emph{do not} behave like real numbers.

There are many excellent resources\cite{knuthArtComputerProgramming1997} on understanding the theory behind \fp{};
we will explain the most essential aspects here.
We represent \fp{} values $r \in \mathbb{R}$ in scientific notation $r = f \times b^{e-q}$,
for some exponent $e$, a fixed value called an \emph{excess} $q$,\footnote{The purpose of the excess $q$ is so that with value of $e > q$ we can represent positive exponents, and $e < q$ we can represent negative exponents.} and fraction $f$.\footnote{Another name for the fraction part commonly seen in the literature is ``mantissa''.
  You might occasionally also see ``characteristic'' referring to the exponent.
  We will avoid those terms however.}

We round and pack $e$ and $f$ into bit fields.
This gives us a compact representation for both very small and very large values.

Things get tricky when we start to do math when there is a big disparity between exponent values.
For example, when adding two numbers $u$ and $v$, if the exponents differ by too much, the result of the addition will simply be the larger value.

% TODO illustrative example: e.g. 1,234,567 + 0.0000001 = 1,234,567

The gradual accumulation of these errors leads to surprising mathematical results!
% FIXME: more examples of a NaN popping up out of nowhere, e.g. with bad addition
